## Generalized Additive (Mixed) Models


```{r GenericFun, echo=FALSE, results='hide'}

## In this function, the gam model is applied indicidually to the list of data frames
library(mgcv)

gam.function <- function(x)  {
  if (nrow(x) < 3) return(" nrow() to small ")
  gam.x <-  gam(form, family=quasipoisson, na.action=na.exclude, data = x)
}

# ## Applied to the list of dataframes, return a list of models
# model.list.DM <- lapply(allDistrictLaggedData, FUN=gam.function) 

```

```{r echo=FALSE, results="hide", message=FALSE}

#### association models. All the models are listed below. 


# 1. meteorology only model - all variables all lags

## the formula
form <- as.formula("count ~ s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4)")

## To the entire dataset, returns the model 
mod.DM <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)


# 2. Only past dengue incidences model - all variables all lags

## the formula

form1 <- as.formula("count ~ s(count1,k=4) + s(count2,k=4) + s(count3,k=4) + s(count4,k=4)")
## To the entire dataset, returns the model 

## Short-term incidences
mod.DDShort <- gam(form1, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)

## Short and long together
form2 <- as.formula("count ~ s(count1,k=4) + s(count2,k=4) + s(count3,k=4) + s(count4,k=4) + s(count5,k=4) + s(count6,k=4) + s(count7,k=4) + s(count8,k=4) + s(count9,k=4) + s(count10,k=4) + s(count11,k=4) + s(count12,k=4) + s(count13,k=4) + s(count14,k=4) + s(count15,k=4) + s(count16,k=4) + s(count17,k=4) + s(count18,k=4) + s(count19,k=4) + s(count20,k=4) + s(count21,k=4) + s(count22,k=4) + s(count23,k=4) + s(count25,k=4) + s(count26,k=4) + s(count27,k=4) + s(count28,k=4) + s(count29,k=4) + s(count30,k=4) ")

## Long-Term 
mod.DDLong <- gam(form2, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)
## The closer analysis shows that for dengue lag 1, lag 18 and lag 23 is the most significant. 

form.Optimal <- as.formula("count ~ s(count1,k=4) + s(count2,k=4) + s(count23,k=4)")

mod.DDOptimal <- gam(form.Optimal, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)






# 3. meteorology and past dengue incidences model - all variables all lags

## the formula
form <- as.formula("count ~ s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count8,k=4) + s(count18,k=4) + s(count23,k=4)")

## To the entire dataset, returns the model 
mod.DMD <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)

form <- as.formula("count ~ s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count2,k=4) + s(count23,k=4)")

## To the entire dataset, returns the model 
mod.DMD <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)



# 4. meteorology, past dengue incidences and surrounding dengue incidences model - all variables all lags

## the formula
form <- as.formula("count ~  s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count2,k=4) + s(count23,k=4) +  s(surroundingCountVec1,k=4) + s(surroundingCountVec2,k=4)")


## To the entire dataset, returns the model 
mod.DMDS_Short <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)


form <- as.formula("count ~  s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count2,k=4) + s(count23,k=4) + +  s(surroundingCountVec1,k=4) + s(surroundingCountVec2,k=4) +  s(surroundingCountVec12,k=4)")

mod.DMDS_Optimal <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)


# 5. Garbage offset along with meteorology, past dengue incidences and surrounding dengue incidences model - all variables all lags

## the formula
form <- as.formula("count ~ s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count2,k=4) + s(count23,k=4) +  s(surroundingCountVec1,k=4) + s(surroundingCountVec2,k=4) +  s(surroundingCountVec12,k=4) + offset(log(garbage)) + s(garbage)")

## To the entire dataset, returns the model 
mod.DMDSO <- gam(form, family=quasipoisson, na.action=na.exclude, data = trainTill2012_df)

```


```{r dSum, echo=FALSE, results="hide"}

d <- matrix(trainTill2012_df$count, nrow = 60, byrow = FALSE)
d <- as.data.frame(d)
dSum <- rowSums(d,na.rm = TRUE)

```


### Meterological Data 

In this model, the association of meterologiocal variables i.e. DTR and averrage monthly rainfall is considered.  I call this **Dengue-Meteorological model**.


The above summary in Appendix \@ref(appDM) suggests that all the temperature and rain lag variables are important factors. Let's visualize the additive model in Figure \@ref(fig:D-M). 

```{r D-M, fig.cap= '**Association between the meteorological variables and dengue over lags of 0-3 months.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', echo=FALSE}

par(mfrow=c(2,4))
plot.gam(mod.DM, ylab="log(RR)")

```

```{r DMPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

par(mfrow=c(1,1))
p <- fitted.values(mod.DM)

a<- predict(mod.DM, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend(x = 20, y = 1400, c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Metereology Optimal")



RMSE_DM <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DM <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DM <-   summary(mod.DM)$r.sq
dev.expl_DM <-  summary(mod.DM)$dev.expl

summary_DM <- data.frame("Meteorology Model", RMSE_DM, SRMSE_DM, r.sq_DM, dev.expl_DM)
names(summary_DM) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```



```{r DMTable, echo=FALSE, comment=NA}
knitr::kable(summary_DM, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Metereology Model.'
)
```


### Dengue Surveillance Data 

In this model the association of past denge incidences is considered. 

#### Short-term Lag Model

The summary of the model is shown in Appendix \@ref(appDDShort). Let's visualize the additive model in Figure \@ref(fig:DDShort). 

```{r DDShort, fig.cap= '**Association between past dengue count over lags of 1-4 months and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', echo=FALSE}

par(mfrow=c(1,4))
plot.gam(mod.DDShort, ylab="log(RR)")

```

```{r DDShortPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DDShort ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Short-term Lag Model")

RMSE_DDShort <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DDShort<-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DDShort <-   summary(model)$r.sq
dev.expl_DDShort <-  summary(model)$dev.expl

summary_DDShort <- data.frame("Short-term Lag Model", RMSE_DDShort, SRMSE_DDShort, r.sq_DDShort, dev.expl_DDShort)
names(summary_DDShort) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DDShortTable, echo=FALSE, comment=NA}
knitr::kable(summary_DDShort, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Short-term Lag Model.'
)
```


#### Long-term Lag Model

```{r AR_GAM, echo=FALSE, results='hide'}
library(dlnm)

totalDHF <- aggregate(count ~ date_sick_year + date_sick_month, trainTill2012_df ,sum)
totalDHF <- totalDHF[order(totalDHF$date_sick_year), ]

## Taking a one lag 

totalDHF_lagged <- aggregate(count1 ~ date_sick_year + date_sick_month, trainTill2012_df ,sum)
totalDHF_lagged <- totalDHF_lagged[order(totalDHF_lagged$date_sick_year), ]
totalDHF_lagged <- c(NA, totalDHF_lagged$count1)



cb1.temp <- crossbasis(totalDHF_lagged, lag=30, argvar=list(df=4), arglag=list(fun="poly",degree=4))
summary(cb1.temp)
mod.dlnm <- glm(totalDHF$count ~  cb1.temp , family=quasipoisson())
pred1.temp <- crosspred(cb1.temp, mod.dlnm)

```

I show the simulated lag–response surfaces as relative risk in Figure \@ref(fig:lag-response). 

```{r lag-response, echo=FALSE, results="hide", fig.cap= 'This shows the relation between the case intensity and dengue incidences at the lag months', fig.align='center',}

plot(pred1.temp, "contour", key.title=title("RR"), plot.title=title("Dengue Lag Influence",xlab="Dengue Counts",ylab="Lag"))


```


The summary of the model is shown in Appendix \@ref(appDDOptimal). Let's visualize the additive model in Figure \@ref(fig:DDOptimal). 

```{r DDOptimal, fig.cap= '**Association between past dengue count over optimal lags within 1-30 months and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', echo=FALSE}

par(mfrow=c(1,3))
plot.gam(mod.DDOptimal, ylab="log(RR)")

```

```{r DDOptimalPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DDOptimal ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Optimal-term Lag Model")

RMSE_DDOptimal  <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DDOptimal <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DDOptimal <-   summary(model)$r.sq
dev.expl_DDOptimal <-  summary(model)$dev.expl

summary_DDOptimal <- data.frame("Optimal-term Lag Model", RMSE_DDOptimal, SRMSE_DDOptimal, r.sq_DDOptimal, dev.expl_DDOptimal)
names(summary_DDOptimal) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DDOptimalTable, echo=FALSE, comment=NA}
knitr::kable(summary_DDOptimal, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Optimal-term Lag Model.'
)
```


### Meteorology and Optimal-term Lag Model.


The summary of the model is shown in Appendix \@ref(appDMD). Let's visualize the additive model in Figure \@ref(fig:DMD). 

```{r DMD, fig.cap= '**Association between the meteorological variables, past dengue count over optimal lags within 1-30 months and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', fig.width = 10, fig.height=10, echo=FALSE}

par(mfrow=c(3,4))
plot.gam(mod.DMD, ylab="log(RR)")

```

```{r DMDPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DMD ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Meteorology and Optimal-term Lag Model")

RMSE_DMD  <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DMD <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DMD <-   summary(model)$r.sq
dev.expl_DMD <-  summary(model)$dev.expl

summary_DMD <- data.frame("Meteorology and Optimal-term Lag Model", RMSE_DMD, SRMSE_DMD, r.sq_DMD, dev.expl_DMD)
names(summary_DMD) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DMDTable, echo=FALSE, comment=NA}
knitr::kable(summary_DMD, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Meteorology and Optimal-term Lag Model.'
)
```


### Surrounding Dengue Data 

```{r AR-Surrounding, echo=FALSE, results="hide"}

totalSDHF <- aggregate(surroundingCountVec0 ~ date_sick_year + date_sick_month, trainTill2012_df ,sum)
totalSDHF <- totalSDHF[order(totalSDHF$date_sick_year), ]

## Taking a one lag 



library(dlnm)

cb1S.temp <- crossbasis(totalSDHF$surroundingCountVec0, lag=30, argvar=list(df=4), arglag=list(fun="poly",degree=4))
summary(cb1S.temp)
mod.dlnm <- glm(totalDHF$count ~  cb1S.temp , family=quasipoisson())
pred1.temp <- crosspred(cb1S.temp, mod.dlnm)


```

I show the simulated lag–response surfaces for **surrounding districts** as relative risk in Figure \@ref(fig:lag-response-surround). 

```{r lag-response-surround, echo=FALSE, results="hide", fig.cap= 'This shows the relation between the case intensity and dengue incidences in surrounding districts at the lag months', fig.align='center',}

plot(pred1.temp, "contour", key.title=title("RR"), 
     plot.title=title("Surrounding Dengue Lag Influence",xlab="Dengue Counts",ylab="Lag"))


```


### Meteorology, Optimal-term and Short-term Surrounding Lag Model

The summary of the model is shown in Appendix \@ref(appDMDS_Short). Let's visualize the additive model in Figure \@ref(fig:DMDS_Short). 

```{r DMDS_Short, fig.cap= '**Association between the meteorological variables, past dengue count over optimal lags within 1-30 months, surroinding district count over 0-3 months and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', fig.width = 10, fig.height=10, echo=FALSE}

par(mfrow=c(4,4))
plot.gam(mod.DMDS_Short, ylab="log(RR)")


```

```{r DMDSShort_Pred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DMDS_Short ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Met, Optimal(D) and Short (D-S) Lag Model")

RMSE_DMDS_Short <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DMDS_Short <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DMDS_Short <-   summary(model)$r.sq
dev.expl_DMDS_Short <-  summary(model)$dev.expl

summary_DMDS_Short <- data.frame("Meteorology, Optimal(D) Short(D-S) Lag Model", RMSE_DMDS_Short, SRMSE_DMDS_Short, r.sq_DMDS_Short, dev.expl_DMDS_Short)
names(summary_DMDS_Short) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DMDShortTable, echo=FALSE, comment=NA}
knitr::kable(summary_DMDS_Short, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Meteorology and Optimal-term Lag Model.'
)
```


### Meteorology, Optimal-term and Optimal-term Surrounding Lag Model

The summary of the model is shown in Appendix \@ref(appDMDS_Optimal). Let's visualize the additive model in Figure \@ref(fig:DMDS_Optimal). 

```{r DMDS_Optimal, fig.cap= '**Association between the meteorological variables, past dengue count over optimal lags within 1-30 months, surrounding district count over 0-30 months and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.align='center', fig.width = 10, fig.height=10, echo=FALSE}

par(mfrow=c(4,4))
plot.gam(mod.DMDS_Optimal, ylab="log(RR)")

```

```{r DMDSOptimalPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DMDS_Optimal ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Meteorology, Optimal (D, D-S) Lag Model")

RMSE_DMDS_Optimal <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DMDS_Optimal <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))
r.sq_DMDS_Optimal <-   summary(model)$r.sq
dev.expl_DMDS_Optimal <-  summary(model)$dev.expl

summary_DMDS_Optimal <- data.frame("Meteorology, Optimal (D, D-S) Lag Model", RMSE_DMDS_Optimal, SRMSE_DMDS_Optimal, r.sq_DMDS_Optimal, dev.expl_DMDS_Optimal)
names(summary_DMDS_Optimal) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DMDSOptiTable, echo=FALSE, comment=NA}
knitr::kable(summary_DMDS_Optimal, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Meteorology and Optimal-term Lag Model.'
)
```


## Social-Economic Data 


The summary of the model is shown in Appendix \@ref(appDMDSO). Let's visualize the additive model in Figure \@ref(fig:DMDSO). 

```{r DMDSO, fig.cap= '**Association between the meteorological variables, past dengue count over optimal lags within 1-30 months, surrounding district count over 0-30 months, garbage data and the dengue outbreak.**. Solid lines represent relative risks (RR) of dengue cases and dottted lines depict the upper and lower limits of 95% confidence intervals.', fig.width = 10, fig.height=10, fig.align='center', echo=FALSE}

par(mfrow=c(4,4))
plot.gam(mod.DMDSO, ylab="log(RR)")

```

```{r DMDSShortPred, echo=FALSE, fig.cap="Monthly Observed and predicted dengue cases (2008-2012). "}

model <- mod.DMDSO ## name of the model 

par(mfrow=c(1,1))
p <- fitted.values(model)
a <- predict(model, type="response")
a <- matrix(a, nrow = 60, byrow = FALSE)
a <- as.data.frame(a)
aSum <- rowSums(a ,na.rm = TRUE)
is.na(aSum) <- !aSum

plot(1:60, dSum, type="l",ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum,type="l", col="red")
legend('top', c("Observed", "Predicted"),lty=1, col=c('black', 'red'), bty='n', cex=.75)

title(main="Social-economic data Included")

RMSE_DMDSO<-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))
SRMSE_DMDSO <-  sqrt(mean((trainTill2012_df$count-p)^2,na.rm=T))/sqrt(mean((trainTill2012_df$count)^2,na.rm=T))


r.sq_DMDSO <-   summary(model)$r.sq
dev.expl_DMDSO <-  summary(model)$dev.expl

summary_DMDSO<- data.frame("Social-economic data Included", RMSE_DMDSO, SRMSE_DMDSO, r.sq_DMDSO, dev.expl_DMDSO)
names(summary_DMDSO) <- c("Model Name", "RMSE", "SRMSE", "R-sq.(adj)", "Deviance Explained")

```


```{r DMDSOTable, echo=FALSE, comment=NA}
knitr::kable(summary_DMDSO, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of Social-economic data Included.'
)
```



## Predictive Performance Statistics 

On the **training dataset**.

```{r TotalPerformance, echo=FALSE, comment=NA}

totalPerformance <- rbind(summary_DM, summary_DDShort, summary_DDOptimal, summary_DMD, summary_DMDS_Short, summary_DMDS_Optimal, summary_DMDSO)

knitr::kable(totalPerformance, booktabs = TRUE,
  caption = 'Predictive Performance Statistics of All Models'
)
```


## Evaluation 


 <!-- WholeDataLagged -->
 
<!-- Formula: -->
<!-- count ~ s(templ0, k = 4) + s(templ1, k = 4) + s(templ2, k = 4) +  -->
<!--     s(templ3, k = 4) + s(rainl0, k = 4) + s(rainl1, k = 4) +  -->
<!--     s(rainl2, k = 4) + s(rainl3, k = 4) + s(count1, k = 4) +  -->
<!--     s(count8, k = 4) + s(count18, k = 4) + s(count23, k = 4) +  -->
<!--     s(count25, k = 4) + s(surroundingCountVec0, k = 4) + s(surroundingCountVec13,  -->
<!--     k = 4) + s(surroundingCountVec14, k = 4) + s(surroundingCountVec15,  -->
<!--     k = 4) -->



```{r validation, echo=FALSE}

form <- as.formula("count ~  s(templ0, k=4) + s(templ1, k=4) + s(templ2,k=4) + s(templ3,k=4) + s(rainl0,k=4) + s(rainl1,k=4) + s(rainl2,k=4) + s(rainl3,k=4) + s(count1,k=4) + s(count2,k=4) + s(count23,k=4) +  s(surroundingCountVec1,k=4) + s(surroundingCountVec2,k=4) +  s(surroundingCountVec12,k=4) ")

preddata <- subset(WholeDataLagged, select = c(templ0, templ1, templ2, templ3, rainl0, rainl1, rainl2, rainl3, count1, count2,  count23, surroundingCountVec1, surroundingCountVec2, surroundingCountVec12))


train1 <- subset(WholeDataLagged,date_sick_year < 2013)
train2 <- subset(WholeDataLagged,date_sick_year < 2014)
train3 <- subset(WholeDataLagged,date_sick_year < 2015)

## for Train 1
mod.DMDS_1 <- gam(form, family=quasipoisson, na.action=na.exclude, data = train1)
mod.DMDS_2 <- gam(form, family=quasipoisson, na.action=na.exclude, data = train2)
mod.DMDS_3 <- gam(form, family=quasipoisson, na.action=na.exclude, data = train3)

WholeDataLagged$predict1 <-predict(mod.DMDS_1, type="response", newdata=preddata)
WholeDataLagged$predict2 <-predict(mod.DMDS_2, type="response", newdata=preddata)
WholeDataLagged$predict3 <-predict(mod.DMDS_3, type="response", newdata=preddata)

train1 <- subset(WholeDataLagged,date_sick_year < 2013)
train2 <- subset(WholeDataLagged,date_sick_year < 2014)
train3 <- subset(WholeDataLagged,date_sick_year < 2015)

pred1 <- subset(WholeDataLagged,date_sick_year> 2012)
pred2 <- subset(WholeDataLagged,date_sick_year> 2013)
pred3 <- subset(WholeDataLagged,date_sick_year> 2014)
pred4 <- subset(WholeDataLagged,date_sick_year> 2012 & date_sick_year < 2014)
pred5 <- subset(WholeDataLagged,date_sick_year> 2013 & date_sick_year < 2015)


p1 <- WholeDataLagged$predict1
a <- matrix(p1, nrow = 96, byrow = FALSE)
a <- as.data.frame(a)
aSum_1 <- rowSums(a ,na.rm = TRUE)
is.na(aSum_1) <- !aSum_1

d <- matrix(WholeDataLagged$count, nrow = 96, byrow = FALSE)
d <- as.data.frame(d)
dSum <- rowSums(d,na.rm = TRUE)

par(mfrow=c(1,1))
plot(1:96, dSum, type="l", ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum_1 [1:60],type="l", col="red")
points(61:96, aSum_1[61:96],type="l", col="blue")
abline(h=790, col = "gray60")
legend("topleft", c("Observed", "Predicted on Training Set", "Predicted on Validation Set"),lty=1, col=c('black', 'red', "blue"), bty='n', cex=.75)

abline(h=790, col = "gray60")
#title("Training Dataset (2008-2012) was used.")


#for training data

aa <- "2008-2012"
  
a <-sqrt(mean((train1$count-train1$predict1)^2,na.rm=T))/sqrt(mean((train1$count)^2,na.rm=T))

#for validation data 2013-2015
b<-sqrt(mean((pred1$count-pred1$predict1)^2,na.rm=T))/sqrt(mean((pred1$count)^2,na.rm=T))

#for validation data 2014-2015
c<-sqrt(mean((pred2$count-pred2$predict1)^2,na.rm=T))/sqrt(mean((pred2$count)^2,na.rm=T))

#for validation data Only 2013
d<-sqrt(mean((pred4$count-pred4$predict1)^2,na.rm=T))/sqrt(mean((pred4$count)^2,na.rm=T))

#for validation data Only 2015
dd<-sqrt(mean((pred5$count-pred5$predict1)^2,na.rm=T))/sqrt(mean((pred5$count)^2,na.rm=T))

#for validation data Only 2015
e<-sqrt(mean((pred3$count-pred3$predict1)^2,na.rm=T))/sqrt(mean((pred3$count)^2,na.rm=T))


pred_accuracy1 <- data.frame(aa, a, b, c, d,dd, e)
names(pred_accuracy1) <- c("Training Dataset", "In-sample Error", "Out-Sample (2013-2015)", "Out-Sample (2014-2015)", "Out-Sample (2013)", "Out-Sample (2014)", "Out-Sample (2015)" )


## for train 2

p2 <- WholeDataLagged$predict2
a <- matrix(p2, nrow = 96, byrow = FALSE)
a <- as.data.frame(a)
aSum_2 <- rowSums(a ,na.rm = TRUE)
is.na(aSum_2) <- !aSum_2

par(mfrow=c(1,1))
plot(1:96, dSum, type="l", ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum_2[1:72],type="l", col="red")
points(73:96, aSum_2[73:96],type="l", col="blue")
legend('topleft', c("Observed", "Predicted on Training Set", "Predicted on Validation Set"),lty=1, col=c('black', 'red', "blue"), bty='n', cex=.75)

abline(h=790, col = "gray60")
#title("Training Dataset (2008-2013) was used.")


aa <- "2008-2013"
  
  
#for training data
a <- sqrt(mean((train2$count-train2$predict2)^2,na.rm=T))/sqrt(mean((train2$count)^2,na.rm=T))

#for validation data 2014-2015
c <- sqrt(mean((pred2$count-pred2$predict2)^2,na.rm=T))/sqrt(mean((pred2$count)^2,na.rm=T))

#for validation data Only 2015
e <- sqrt(mean((pred3$count-pred3$predict2)^2,na.rm=T))/sqrt(mean((pred3$count)^2,na.rm=T))

#for validation data Only 2014
dd <- sqrt(mean((pred4$count-pred4$predict2)^2,na.rm=T))/sqrt(mean((pred4$count)^2,na.rm=T))


pred_accuracy2 <- data.frame(aa, a, b = NA, c, d = NA,dd, e)
names(pred_accuracy2) <- c("Training Dataset", "In-sample Error", "Out-Sample (2013-2015)", "Out-Sample (2014-2015)", "Out-Sample (2013)", "Out-Sample (2014)", "Out-Sample (2015)" )


## for train 3

p3 <- WholeDataLagged$predict3
a <- matrix(p3, nrow = 96, byrow = FALSE)
a <- as.data.frame(a)
aSum_3 <- rowSums(a ,na.rm = TRUE)
is.na(aSum_3) <- !aSum_3

par(mfrow=c(1,1))
plot(1:96, dSum, type="l", ylab="Dengue Cases",axes=T,xlab="Months")
points(aSum_3[1:84],type="l", col="red")
points(85:96, aSum_3[85:96],type="l", col="blue")
abline(h=790, col = "gray60")
legend('topleft', c("Observed", "Predicted on Training Set", "Predicted on Validation Set"),lty=1, col=c('black', 'red', "blue"), bty='n', cex=.75)

abline(h=790, col = "gray60")
#title("Training Dataset (2008-2014) was used.")

aa <- "2008-2014"

#for training data
a <- sqrt(mean((train3$count-train3$predict3)^2,na.rm=T))/sqrt(mean((train3$count)^2,na.rm=T))

#for validation data Only 2015
e <-  sqrt(mean((pred3$count-pred3$predict3)^2,na.rm=T))/sqrt(mean((pred3$count)^2,na.rm=T))

pred_accuracy3 <- data.frame(aa, a , b = NA, c = NA, d = NA,dd = NA, e)
names(pred_accuracy3) <- c("Training Dataset", "In-sample Error", "Out-Sample (2013-2015)", "Out-Sample (2014-2015)", "Out-Sample (2013)", "Out-Sample (2014)", "Out-Sample (2015)" )



```


```{r totalAccuracy, echo=FALSE}

all <- rbind(pred_accuracy1, pred_accuracy2, pred_accuracy3)
knitr::kable(all, booktabs = TRUE,
  caption = 'Predictive Performance Statistics measured using SRMSE'
)


```


 


