# (APPENDIX) Appendix {-} 

# Software Tools


```{r LoadPackagesAgain, message=FALSE}

require(ggplot2) # used in plotting 
require(reshape)
require(plotly)
require(bnlearn) # used for Bayesian Networks
require(mgcv)    # used for Generalized Additive Modeling 
require(fpp)
require(lubridate)
require(bsts)
require(dplyr)
require(CausalImpact)
require(xtable)
require(dplyr)
require(tidyr)
require(htmlwidgets)
require(DT)
```

# Supporting information 

## Cross-correlation between Outcome and Predictors

```{r Cross-Cor, echo=FALSE, results="hide", message=FALSE}

attach(training2008_2012)

lag <- as.integer(c(0:3))
par(mfrow=c(2,2))
#ylim=c(0, 1)


# dengue 
short_den <- data.frame(count0, count1, count2, count3)
cor(count, short_den, method="spearman", use="pairwise.complete.obs")
plot(lag, as.numeric(cor(count, short_den , method="spearman", use="pairwise.complete.obs")), ylab="Correlation coefficient", xlab="Lag of monthly dengue cases")          
title(main="Dengue cases vs. Dengue cases")


# temp
short_temp <- data.frame(templ0, templ1, templ2, templ3)
cor(count, short_temp, method="spearman", use="pairwise.complete.obs")
plot(lag, as.numeric(cor(count, short_temp, method="spearman", use="pairwise.complete.obs")), ylab="Correlation coefficient", xlab="Lag of monthly temperature")          
title(main="Dengue cases vs. Temperature")

# rain
short_rain <- data.frame(rainl0, rainl1, rainl2, rainl3)
cor(count, short_rain, method="spearman", use="pairwise.complete.obs")
plot(lag, as.numeric(cor(count, short_rain, method="spearman", use="pairwise.complete.obs")), ylab="Correlation coefficient", xlab="Lag of monthly rainfall")          
title(main="Dengue cases vs. Rainfall")


# with neighboring districts

short_neighbor <- data.frame(surroundingCountVec0, surroundingCountVec1, surroundingCountVec2, surroundingCountVec3)
cor(count, short_neighbor, method="spearman", use="pairwise.complete.obs")
plot(lag, as.numeric(cor(count, short_neighbor , method="spearman", use="pairwise.complete.obs")), ylab="Correlation coefficient", xlab="Lag of monthly dengue cases")          
title(main="Dengue cases vs. Surrounding Dengue cases")


detach(training2008_2012)


```


## Summary of Meteorology Model {#appDM}

```{r Summary-D-M, echo=FALSE, comment=NA}
summary(mod.DM)
```


## Summary of  Short-term Lag Model {#appDDShort}

```{r Summary-DDShort, echo=FALSE, comment=NA}
summary(mod.DDShort)
```

## Summary of Long-term Lag Model {#appDDLong}

```{r Summary-DDLong, echo=FALSE, comment=NA}
summary(mod.DDLong)
```

## Summary of Optimal-term Lag Model {#appDDOptimal}

```{r Summary-DDOptimal, echo=FALSE, comment=NA}
summary(mod.DDOptimal)
```

## Summary of Meteorology and Optimal-term Lag Model {#appDMD}

```{r Summary-DMD, echo=FALSE, comment=NA}
summary(mod.DMD)
```

## Summary of Meteorology, Optimal-term and Short-term Surrounding Lag Model {#appDMDS_Optimal}

```{r DMDS_ShortC, echo=FALSE, comment=NA}
summary(mod.DMDS_Short)
```

## Meteorology, Optimal-term and Optimal-term Surrounding Lag Model

```{r Summary_DDOptimal, echo=FALSE, comment=NA}
summary(mod.DMDS_Optimal)
```

## Social Data Included {#appDMDSO}

```{r SummaryDMDSO, echo=FALSE, comment=NA}
summary(mod.DMDSO)
```

## Initial Analyses Summary







